---
title: "06-takehomequiz"
author: "Milo Coolman"
format: 
  html:
    embed-resources: true
---

All work presented is my own, and I have followed all rules for collaboration. If I used any AI on this assignment, I have clearly stated both (1) what my search prompt was and (2) what I used from the AI answer.
- Milo Coolman

```{r}
library(tidyverse)
library(here)
attendance_full <- read_csv(here("data/attendance_113.csv"))
attendance_full
```

# Question 1
```{r}
attendance_q1 <- attendance_full |> group_by(section) |> 
  pivot_longer(cols = c(3:40), names_to = "class", values_to = "attendance") |>
  mutate(attendance_chr = case_when(attendance == "x" ~ "1",
                                     attendance = is.na(attendance) ~ "0")) |>
  mutate(attendance_dbl = as.double(attendance_chr)) |>
  ungroup() |>
  group_by(section, class) |>
  summarise(mean_attendance = sum(attendance_dbl))

attendance_q1.2 <- attendance_q1 |> ungroup() |> group_by(section) |>
  summarise(mean_section = mean(mean_attendance))

ggplot(data = attendance_q1, aes(x = section, y = mean_attendance)) +
  geom_point()
  geom_point(data = attendance_q1.2, aes(x = section, y = mean_section),
             colour = "purple", shape = 8) +
  scale_x_continuous(breaks = c(810, 920)) +
```

# Question 2
I do not believe that there is evidence that the number of absences was substantially different for the two sections as the mean absences was roughly 2.5 for both.

# Question 3
```{r}
attendance_full |> group_by(section) |> 
  pivot_longer(cols = c(3:40), names_to = "class", values_to = "attendance") |>
  mutate(attendance_chr = case_when(attendance == "x" ~ "1",
                                     attendance = is.na(attendance) ~ "0")) |>
  mutate(attendance_dbl = as.double(attendance_chr)) |>
  ungroup() |>
  group_by(section, class) |>
  summarise(mean_attendance = sum(attendance_dbl)) |>
  filter(mean_attendance == max(mean_attendance))
```
The Friday on Week 12 is the first day of Thanksgiving, so I would guess that many people especially those with long travels back home would have been willing to skip that class so as to get home faster.


